Active Object Search with a Mobile Device\\for People with Visual Impairments

### 2018-12-28 review_edition/reviewed.tex

\noindent To evaluate our system, we designed a set of experiments that determined how effective the MDP and its policies are at guiding the user in an object search task with the smartphone's camera.
% This section discusses the experiments and their results.

\subsection{Experiment Design}

\noindent For the experiment, the MDP policies were integrated into an Android application that uses the camera to provide observation data and track the pose and viewing direction. We avoided adding unnecessary variability and complexity to the experiment by providing visual waypoints on screen which the participants were allowed to use. A generic environment mimicking a typical office desk layout was set up as the experiment environment and contains 7 different objects (i.e.\ encoded QR codes), one of which could be selected for each experiment run. See Figure~\ref{fig:env-picture} for a picture of the environment.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{../figures/test_env_picture.jpg}
  \caption{A picture of the environment used for the experiments. Each QR code represents an object.}\label{fig:env-picture}
\end{figure}

For each experiment, the participant was placed roughly \SI{1}{\meter} from the barcodes and was asked not to move from that spot during the experiment. The participant then pressed a button on the app that randomly selected a target object and was then guided towards that target by the app. Since the participants were allowed to use their vision, the targets were randomly selected and the participants were uninformed of what the target object was until it was found. This was done to avoid the participants learning the target objects' locations between each search run.

During in-house tests, we noted that the system sometimes makes guidance mistakes that lead the user towards an uncluttered edge of the search space, at which point the system had difficulty guiding the user back to the centre. To avoid this having an adverse impact and drawing out the experiments too long, we set a search step limit of 15, which means that the search was terminated when the number of waypoints exceeded 15. A search run therefore ended when the participant either successfully found the target object by pointing the phone camera to it and scanning the barcode, or exceeded the waypoint limit. After this, the participant then resets to the central position, selects a new target object and resumes the experiment.

The participants were left to repeat this process until we recorded 10 searches per object, giving us a total number of 70 recorded search samples per participant. We tested a total of 10 different participants (sex: 8 M, 2 F;\@ age: $\mu$ 31.2 years, $\sigma$ 8 years), most of whom were sourced from our laboratory. This gives our dataset a total of 700 search samples.

@@

\noindent We identified 3 different measures to evaluate the system's performance: target acquisition rate (TAR), number of steps to the target and the total time it took to find a target object. We present and discuss the results for each of these parameters in the next sections. The results for each individual participant is presented in Table~\ref{tab:results-full} and Table~\ref{tab:results-summary} gives the results across all of the participants. To provide a baseline measure for the ideal case, we ran a number of simulations in an environment mimicking the experiment setup with a user who perfectly executes the policy, i.e.\ $u=u^*$. The TAR and steps to target results for the simulation are included in Table~\ref{tab:results-summary}.

@@

\begin{table}[b]                      
  \centering                       
  \caption{The means and standard deviations for the entire participant population for the experiments and simulations.}\label{tab:results-summary}
  \begin{tabular}{ccc}
    \toprule
    \multirow{2}{*}{TAR [\%]} & Experiment & 82 $\pm$ 11  \\ 
			      & Simulation & 99.7 \\ \midrule
    \multirow{2}{*}{Num. Steps} & Experiment & 6.8 $\pm$ 5.1 \\ 
			        & Simulation & 5.4 \\ \midrule
    Time [s] & Experiment & 34 $\pm$ 23  \\ \midrule
    \bottomrule
  \end{tabular}
\end{table}

@@

\noindent The TAR is a measure of how successful the system was at directing a participant to the target object within the 15-step limit we set and is simply calculated as a proportion of the number completed searches vs.\ the total number of searches. Please note, however, that this plot is dependant on the step-limit we selected and should not be taken as an absolute measure of performance, but is rather presented to provide context to the other results we present. Figure~\ref{fig:tar-steps} shows the TAR as a function of the step-limit and it shows that there is a gradual increase in the TAR as the step limit increases, but tapers off as the step-limit increases.

@@

Figure~\ref{fig:tar-objects} shows the TAR for each object in $\mathbf{O}$. There are variations the TAR for the different objects, with the smaller objects typically being the hardest to find. However, the differences are not extreme and indicate that all the objects in $\mathbf{O}$ are roughly equally hard to find. This is also displayed in the simulation's TAR in Table~\ref{tab:results-summary}, which could not achieve 100\% because of the difficulty the agent had in finding the objects on the fringes of the environment.

@@

Failure cases were typically caused by the system entering a no-recovery state where the system directed a user into dead-space with no spatial information (e.g.\ a ceiling or wall section) and could not gather the information it required to intelligently guide the user. Subsequent system updates will have to consider implementing fall-back methods or automatic recovery modes that can detect entry of a no-recovery state (e.g.\ exceeding a set number of steps/time without making an observation) and guide the user back to the initial position and restart a search.

@@

\noindent The number of steps to the target indicates the number of waypoints the system generated for the participant during the guidance process and is a major indication of system performance, where less waypoints means faster target acquisition and better performance. Figure~\ref{fig:nsteps-participants} shows the cumulative distribution of the number of steps to the target for all the participants.

@@

The number of waypoints each participant required is fairly evenly spread across all of the participants, with the majority of searches ending within 4 steps of initialisation or ending. The population mean and standard deviation is 6.8 and 5.1 waypoints respectively. This is a reasonable result, since most target objects were placed within approximately 4 grid squares away from the participants' initial looking directions. However, the relatively high standard deviation should be noted and can also be reduced by implementing the automatic recovery mode mentioned previously, and avoiding information-sparse areas.

@@

\noindent The distribution of the total time it took the participants to reach the target object is given in Figure~\ref{fig:time-participants}. We see that the distribution is heavily skewed to the bottom with a long tail. The mean and standard deviation of the data is 29.8s and 23s respectively. %, is far greater than for the number of waypoints they required. Indeed, there is not a strong relationship between the number of waypoints a participant required and the time they took to find the target, as showcased by participant $s7$'s wide mean waypoint count and spread in Figure~\ref{fig:nsteps-participants} and the small spread and close-to-average time to target in Figure~\ref{fig:time-participants}. Furthermore, the opposite can be observed in participant $s8$'s small spread and low mean waypoint count in Figure~\ref{fig:nsteps-participants} and high spread and high mean time to target in Figure~\ref{fig:time-participants}.

@@

In comparison to the VizWiz system~\cite{bigham2010vizwiz} (mean 92s, standard deviation 37.7s), our results look very reasonable, although there might be significant variations in case of participants with vision impairments.

>>>>

\noindent To evaluate our system, we designed a set of experiments that determined how effective the MDP and its policies are at guiding the user in an object search task with the smartphone's camera. Since the focus of this work is on the algorithm for active object search, and not on the actual human performance, in the following experiments the system was tested by participants without any significant visual impairment. As explained in the previous section, this simplified the experimental design, allowing us to use the smartphone's display for the guidance instructions.
% This section discusses the experiments and their results.

\subsection{Experimental Design}

\noindent For the experiment, the MDP policies were integrated into an Android application that uses the camera to provide observation data and track the pose and viewing direction. Guidance instructions towards the waypoints were visualised on the screen, which the participants were allowed to use. The experimental environment mimicked a typical office desk layout and contained 7 different objects (i.e.\ encoded QR codes), one of which could be selected for each experiment run. See Figure~\ref{fig:env-picture} for a snapshot of the environment.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{../figures/test_env_picture.jpg}
  \caption{A snapshot of the environment used for the experiments. Each QR code represents an object.}\label{fig:env-picture}
\end{figure}

For each experiment, the participant was placed approximately at 1m from the closest barcodes and was asked to remain on that spot during the experiment. The participant started by pressing a button on the app, which randomly selected a target object and then guided the user towards it. Since the participants were allowed to use the smartphone's display, the target was randomly selected by the app without informing them, at least until it was found. This prevented the participants from learning the target objects' locations between subsequent runs of the experiment.

To avoid pointing at uncluttered edges of the search space, where the system had difficulty guiding the user back to the centre, we set a search step-limit of 15, which means the search was terminated when the number of waypoints generated by the system exceeded 15. A search run therefore ended when the participant either successfully found the target object by pointing the phone camera to it and scanning the barcode, or exceeded the waypoint limit. After this, the participant then restarted from the central position, generating a new random target object and repeating the experiment.

We recorded 10 searches per object, giving us a total number of 70 search samples per participant. The system was tested by 10 participants, mostly from our research group. Our final dataset consisted therefore of 700 search samples, which are analysed next.

@@

\noindent We identified 3 different measures to evaluate the system's performance: target acquisition rate (TAR), number of steps to the target and the total time it took to find a target object. We present and discuss the results for each of these parameters in the next sections. The results for each individual participant are presented in Table~\ref{tab:results-full}, while those across all the participants are shown in Table~\ref{tab:results-summary}. To provide a baseline measure for the ideal case, we ran a number of simulations in an environment mimicking the experiment setup with a ``virtual'' user who perfectly executes the policy, i.e.\ $u=u^*$. The TAR and steps to target results for the simulation are included in Table~\ref{tab:results-summary}.

@@

\begin{table}[b]                      
  \centering                       
  \caption{The means and standard deviations for the entire participant population for the experiments and simulations.}\label{tab:results-summary}
  \begin{tabular}{lcc}
%     \toprule
%     \multirow{2}{*}{TAR [\%]} & Experiment & 82 $\pm$ 11  \\ 
% 			      & Simulation & 99.7 \\ \midrule
%     \multirow{2}{*}{Num. Steps} & Experiment & 6.8 $\pm$ 5.1 \\ 
% 			        & Simulation & 5.4 \\ \midrule
%     Time [s] & Experiment & 34 $\pm$ 23  \\ \midrule
%     \bottomrule
    \toprule
     & Experiments & Simulations \\ \midrule
    TAR [\%] & 82 $\pm$ 11 & 99.7 \\ \midrule
    Num. Steps & 6.8 $\pm$ 5.1 & 5.4 \\ \midrule
    Time [s] & 34 $\pm$ 23 & --  \\ \midrule
    \bottomrule
  \end{tabular}
\end{table}

@@

\noindent The TAR is a measure of how successful the system was at directing a participant to the target object within our 15-step limit. It is simply calculated as a ratio between the number of completed searches vs.\ the total number of searches. Please note, however, that this ratio depends also on the step-limit and should not be taken as an absolute measure of performance (i.e. if the step limit was much bigger, the TAR would tend to 100\%). Figure~\ref{fig:tar-steps} shows the TAR as a function of the step-limit and it shows that there is a gradual increase in the TAR as the step limit increases, but tapers off as the step-limit increases.

@@

Figure~\ref{fig:tar-objects} shows the TAR for each object in our set~$\mathbf{O}$. There are TAR variations for the different objects, with the smaller objects typically being the hardest to find. However, the differences are not extreme and indicate that all the objects in $\mathbf{O}$ are roughly equally hard to locate. This is also displayed in the simulation's TAR in Table~\ref{tab:results-summary}, which could not achieve 100\% because of the difficulty the agent had in finding the objects on the fringes of the environment.

@@

Failure cases were typically caused by the system entering a no-recovery state where the user was directed into dead-space with no spatial information (e.g.\ ceiling or wall section). In this case the system could not observe useful clues to intelligently guide the user. Possible improvements for future versions of the algorithm would be to implement some fall-back method that can detect a no-recovery state (e.g.\ exceeding a set number of steps/time without any new object observation) and guide the user back to a position where to restart the search.

@@

\noindent The number of steps to the target indicates the number of waypoints the system generated for the participant during the guidance process. This is a good indication of system performance, where less waypoints means faster target acquisition and therefore better object search strategy. Figure~\ref{fig:nsteps-participants} shows the cumulative distribution of the number of steps to the target for all the participants.

@@

The number of waypoints each participant required is fairly evenly spread across all of the participants, with the majority of searches ending within a few search steps. The population mean and standard deviation is 6.8 and 5.1 waypoints, respectively. This is a reasonable result, since most target objects were placed within approximately 4 grid squares away from the participants' initial looking directions. The relatively high standard deviation is due to the aforementioned no-recovery states and could be reduced by opportune mitigation strategies to avoid uninformative areas.

@@

\noindent The cumulative distribution of the time it took the participants to reach the target object is given in Figure~\ref{fig:time-participants}. We see that the distribution is heavily skewed to the bottom with a long tail. The mean and standard deviation of the data is 29.8s and 23s respectively. %, is far greater than for the number of waypoints they required. Indeed, there is not a strong relationship between the number of waypoints a participant required and the time they took to find the target, as showcased by participant $s7$'s wide mean waypoint count and spread in Figure~\ref{fig:nsteps-participants} and the small spread and close-to-average time to target in Figure~\ref{fig:time-participants}. Furthermore, the opposite can be observed in participant $s8$'s small spread and low mean waypoint count in Figure~\ref{fig:nsteps-participants} and high spread and high mean time to target in Figure~\ref{fig:time-participants}.

@@

In comparison to the remotely-assisted VizWiz system~\cite{bigham2010vizwiz} covered in Section~\ref{sec:previous-work} (mean 92s, standard deviation 37.7s), our results look very encouraging, although there might be significant variations in case of participants with actual vision impairments.
